---
title: "p8105_HW3_rat2134"
author: "Robert Tumasian"
date: "10/8/2019"
output: github_document
---

# Homework 3
```{r, message=FALSE}
#Load required packages
library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
```

## Problem 1
```{r}
#Load dataset
library(p8105.datasets)
data("instacart")
```

The `instacart` dataset contains information on grocery purchases of `r length(unique(pull(instacart, user_id)))` unique online users. The dataset contains `r dim(instacart)[1]` products from a particular order and `r dim(instacart)[2]` variables:

* Order identifier (`order_id`, integer vector)
* Product identifier (`product_id`, integer vector)
* Order each product was added to cart (`add_to_cart_order`, integer vector)
* Ordered previously by user (`reordered`, integer vector) [1 = Yes, 0 = No]
* Customer identifier (`user_id`, integer vector)
* Evaluation set in which the observation is located (`eval_set`, character vector) [all observations in this dataset are from the training set]
* Order sequence number (`order_number`, integer vector) [1 = first order placed by user, n = n'th order placed by user]
* Day of the week the order was placed (`order_dow`, integer vector) [0 = Sunday, 6 = Saturday]
* Hour of the day the order was placed (`order_hour_of_day`, integer vector) [0 = 12 - 12:59 am, 23 = 11 - 11:59 pm]
* Days since last order (`days_since_prior_order`, integer vector) [If user only has one order recorded, `days_since_prior_order` = `NA`]
* Product name (`product_name`, character vector)
* Aisle identifier (`aisle_id`, integer vector)
* Department identifier (`department_id`, integer vector)
* Aisle name (`aisle`, character vector)
* Department name (`department`, character vector)

We can interpret each row of the `instacart` dataset. For instance, consider the first row of the dataset. We can see that order #`r pull(instacart[1,], order_number)` of user #`r pull(instacart[1,], user_id)` was completed on a Wednesday (`order_dow` = `r pull(instacart[1,], order_dow)`) between 10 am and 11 am (`order_hour_of_day` = `r pull(instacart[1,], order_hour_of_day)`). Item #`r pull(instacart[1,],add_to_cart_order)` added to this user's order was Bulgarian Yogurt (`product_id` = #`r pull(instacart[1,], product_id)`), from the yogurt aisle (aisle #`r pull(instacart[1,], aisle_id)`) in the '`r pull(instacart[1,], department)`' department (`department_id` = #`r pull(instacart[1,], department_id)`). This order was completed `r pull(instacart[1,], days_since_prior_order)` days after the user's previous order (order #`r pull(instacart[1,], order_number-1)`). This user has ordered this product previously (`reordered` = `r pull(instacart[1,], reordered)`).

We can interpret each remaining row of the dataset in a similar way to obtain further information regarding the purchases of users over time.

There are a total of `r length(unique(pull(instacart, aisle_id)))` aisles. The top three aisles in which the most items are ordered from are:

1. Fresh vegetables
2. Fresh fruits
3. Packaged vegetables fruits

```{r}
#Top three aisles where most items are ordered from
instacart %>%
  group_by(aisle) %>%
  count(aisle) %>%
  arrange(-n) %>%
  .[1:3,] %>%
  rename("Aisle" = aisle,
         "Number of Items Ordered" = n) %>%
  knitr::kable()
```

```{r}
#Plot showing the number of items ordered from each aisle
instacart %>%
  group_by(aisle, department) %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  ggplot(aes(x = reorder(aisle, -n), y = n, fill = department)) +
    geom_bar(stat="identity") +
    labs(
      title = "Number of items ordered from each aisle, by department",
      x = "Aisle",
      y = "Number of Items Ordered",
      fill = "Department") +
    coord_flip()
```

Here, we can also see that the most items are ordered from the aisles listed above, and these aisles are part of the 'produce' and 'dairy eggs' departments.

```{r}
#Table showing the three most popular items in the 'baking ingredients' aisle
BAKING_INGRED_three_most_popular_items = 
  instacart %>%
    filter(aisle == "baking ingredients") %>%
    select(c(aisle, product_name)) %>%
    count(aisle, product_name) %>%
    arrange(-n) %>%
    .[1:3,] %>%
    rename("Aisle" = aisle,
           "Product Name" = product_name,
           "Number Ordered" = n)

#Table showing the three most popular items in the 'dog food care' aisle
DOG_FOOD_CARE_three_most_popular_items = 
  instacart %>%
    filter(aisle == "dog food care") %>%
    select(c(aisle, product_name)) %>%
    count(aisle, product_name) %>%
    arrange(-n) %>%
    .[1:3,] %>%
    rename("Aisle" = aisle,
           "Product Name" = product_name,
           "Number Ordered" = n)

#Table showing the three most popular items in the 'dog food care' aisle
PACKAGED_VEG_FRUITS_three_most_popular_items = 
  instacart %>%
    filter(aisle == "packaged vegetables fruits") %>%
    select(c(aisle, product_name)) %>%
    count(aisle, product_name) %>%
    arrange(-n) %>%
    .[1:3,] %>%
    rename("Aisle" = aisle,
           "Product Name" = product_name,
           "Number Ordered" = n)

#Merging the three tables above
Three_most_popular_items_in_aisles = 
  bind_rows(BAKING_INGRED_three_most_popular_items,
            DOG_FOOD_CARE_three_most_popular_items,
            PACKAGED_VEG_FRUITS_three_most_popular_items)

knitr::kable(Three_most_popular_items_in_aisles)
```

```{r, warning=FALSE}
#Table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
instacart %>%
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  select(c(order_hour_of_day, order_dow, product_name)) %>%
  group_by(order_dow, product_name) %>%
  summarize(round(mean(order_hour_of_day), digits = 2)) %>%
  pivot_wider(names_from = product_name,
              values_from = 'round(mean(order_hour_of_day), digits = 2)') %>%
  ungroup() %>%
  mutate(order_dow =
           recode(order_dow,
                  "0" = "Sunday", "1" = "Monday", "2" = "Tuesday", 
                  "3" = "Wednesday","4" = "Thursday", "5" = "Friday", 
                  "6" = "Saturday")) %>%
  rename("Day of Week" = order_dow) %>%
  knitr::kable()
```

We can see that, on average, users order both Coffee Ice Cream and Pink Lady Apples at the latest time of the day on Wednesdays between 3-4 pm and 1-2 pm, respectively. In addition, on average, users order Coffee Ice Cream and Pink Lady Apples at the earliest time of the day on Fridays between 10-11 am and Saturdays between 11am-12 pm, respectively.

## Problem 2
























